---
title: "p8105_hw6_qc2337"
output: github_document
---

```{r}
library(tidyverse)
library(p8105.datasets)
library(ggplot2)
set.seed(1)
```


# Problem 2
```{r}
 homicide_df =  
   read_csv("data/homicide_data.csv", na = c("NA","Unknown",".","")) |> 
   janitor::clean_names()
```

create "city_state" variable, filter specific cities and restrict the race, edit the age type.
```{r}
 homicide_clean_df = 
  homicide_df |> 
  mutate(
    city_state = paste(city, state, sep = ", "),
    solved = ifelse(disposition == "Closed by arrest", 1, 0),
    victim_age = as.numeric(victim_age) 
  ) |> 
  filter(
    !city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL"),  
    victim_race %in% c("White", "Black")  
  )
```

fit a logistic regression for city of Baltimore
```{r}
baltimore_df = 
  homicide_clean_df |> 
  filter(city_state == "Baltimore, MD")

fit_logistic = glm(
  solved ~ victim_age + victim_sex + victim_race, data = baltimore_df, family = binomial())

fit_logistic |> 
  broom::tidy(conf.int = TRUE) |> 
  filter(term == "victim_sexMale") |> 
  mutate(
    OR = exp(estimate),
    CI_low = exp(conf.low),
    CI_high = exp(conf.high)
     ) |> 
  select(OR, CI_low, CI_high) |> 
  knitr::kable(digits = 3)
```
From the table we know that the adjusted odds ratio is 0.426, the 95%CI for solving homicides comparing male victims to female victims lies between (0.324, 0.558).


run GLM for each city.
```{r}
fit_logistic_eachcity = function(data) {
  glm_results = glm(solved ~ victim_age + victim_sex + victim_race, data = data, family = binomial())  |> 
    broom::tidy(conf.int = TRUE) |> 
    filter(term == "victim_sexMale") |> 
    mutate(
     OR = exp(estimate),
     CI_low = exp(conf.low),
     CI_high = exp(conf.high)
   )
   
    data.frame(
    city_state = unique(data$city_state),
    OR = glm_results$OR,
    CI_low = glm_results$CI_low,
    CI_high = glm_results$CI_high
  )
}
  
results = homicide_clean_df |> 
  group_by(city_state) |> 
  group_split() |> 
  map_df(fit_logistic_eachcity) 

print(results)
```


Create a plot
```{r}
results_plot = results |> 
  arrange(OR) |> 
  mutate(city_state = fct_inorder(city_state))

results_plot |> 
  ggplot(aes(x = city_state, y = OR)) +
  geom_point() + 
  geom_errorbar(aes(ymin = CI_low, ymax = CI_high), width = 0.2) +
  coord_flip() + 
  labs(
    title = "Estimated Odds Ratios and Confidence Intervals by City",
    x = "City",
    y = "Odds Ratio"
  ) +
  theme_minimal()

```
Comment: Those cities with OR larger than 1, indicates that no significant difference in solving rates between male and female victims.The Albuguergue, NM has the highest OR and widest 95% CI.Other cities have an apparenthigh OR>1 are Stockton,CA and Fresno,CA.

# Problem 3
```{r}
 homicide_df =  
   read_csv("data/birthweight.csv", na = c("NA",".","")) |> 
   janitor::clean_names()
```







