---
title: "p8105_hw6_qc2337"
output: github_document
---

```{r}
library(tidyverse)
library(p8105.datasets)
library(ggplot2)
library(modelr)
library(purrr)
set.seed(1)

```

# Problem 1
```{r}
library(p8105.datasets)
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

fit a model
```{r}
model = lm(tmax ~ tmin, data = weather_df )
```

# Problem 2
```{r}
 homicide_df =  
   read_csv("data/homicide_data.csv", na = c("NA","Unknown",".","")) |> 
   janitor::clean_names()
```

create "city_state" variable, filter specific cities and restrict the race, edit the age type.
```{r}
 homicide_clean_df = 
  homicide_df |> 
  mutate(
    city_state = paste(city, state, sep = ", "),
    solved = ifelse(disposition == "Closed by arrest", 1, 0),
    victim_age = as.numeric(victim_age) 
  ) |> 
  filter(
    !city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL"),  
    victim_race %in% c("White", "Black")  
  )
```

fit a logistic regression for city of Baltimore
```{r}
baltimore_df = 
  homicide_clean_df |> 
  filter(city_state == "Baltimore, MD")

fit_logistic = glm(
  solved ~ victim_age + victim_sex + victim_race, data = baltimore_df, family = binomial())

fit_logistic |> 
  broom::tidy(conf.int = TRUE) |> 
  filter(term == "victim_sexMale") |> 
  mutate(
    OR = exp(estimate),
    CI_low = exp(conf.low),
    CI_high = exp(conf.high)
     ) |> 
  select(OR, CI_low, CI_high) |> 
  knitr::kable(digits = 3)
```
From the table we know that the adjusted odds ratio is 0.426, the 95%CI for solving homicides comparing male victims to female victims lies between (0.324, 0.558).


run GLM for each city.
```{r}
fit_logistic_eachcity = function(data) {
  glm_results = glm(solved ~ victim_age + victim_sex + victim_race, data = data, family = binomial())  |> 
    broom::tidy(conf.int = TRUE) |> 
    filter(term == "victim_sexMale") |> 
    mutate(
     OR = exp(estimate),
     CI_low = exp(conf.low),
     CI_high = exp(conf.high)
   )
   
    data.frame(
    city_state = unique(data$city_state),
    OR = glm_results$OR,
    CI_low = glm_results$CI_low,
    CI_high = glm_results$CI_high
  )
}
  
results = homicide_clean_df |> 
  group_by(city_state) |> 
  group_split() |> 
  map_df(fit_logistic_eachcity) 

print(results)
```


Create a plot
```{r}
results_plot = results |> 
  arrange(OR) |> 
  mutate(city_state = fct_inorder(city_state))

results_plot |> 
  ggplot(aes(x = city_state, y = OR)) +
  geom_point() + 
  geom_errorbar(aes(ymin = CI_low, ymax = CI_high), width = 0.2) +
  coord_flip() + 
  labs(
    title = "Estimated Odds Ratios and Confidence Intervals by City",
    x = "City",
    y = "Odds Ratio"
  ) +
  theme_minimal()

```
Comment: Those cities with OR larger than 1, indicates that no significant difference in solving rates between male and female victims.The Albuguergue, NM has the highest OR and widest 95% CI. Other cities have an apparenthigh OR>1 are Stockton,CA and Fresno,CA.

# Problem 3
```{r}
 birthweight_df =  
   read_csv("data/birthweight.csv", na = c("NA",".","")) |> 
   janitor::clean_names() |> 
   mutate(
     babysex = factor(
       babysex, levels = c(1, 2), 
       labels = c("Male", "Female")),
     frace = factor(
       frace, levels = c(1, 2, 3, 4, 8, 9), 
       labels = c("White", "Black", "Asian", "Puerto Rican", "Other", "Unknown")),
     mrace = factor(
       mrace, levels = c(1, 2, 3, 4, 8, 9), 
       labels = c("White", "Black", "Asian", "Puerto Rican", "Other", "Unknown")),
     malform = factor(
       malform, levels = c(0, 1), labels = c("Absent", "Present"))
  )
```

Propose a regression model for birthweight
```{r}
reg_model = lm(bwt ~ fincome + gaweeks + ppwt + ppbmi + malform + smoken, data = birthweight_df)

summary(reg_model)

birthweight_model = 
  birthweight_df |>
  add_predictions(reg_model) |>
  add_residuals(reg_model)

ggplot(birthweight_model, aes(x = pred, y = resid))+
  geom_point(alpha = 0.5) + 
  geom_smooth() +
  labs(
       title = "Plot of Residuals vs. Fitted Values",
       x = "Fitted Values",
       y = "Residuals") +
  theme_minimal()
```
Describe modeling process: 
 The model includes family monthly income (fincome), gestational weeks (gaweeks),pre-pregnancy weight (ppwt), mother’s pre-pregnancy BMI (ppbmi) + presence of malformations that could affect weight (malform) and average number of cigarettes smoked per day during pregnancy (smoken) as predictors. These choices are based on both social economic status and biological relevance to birthweight.
I also checked the model summary to see the significance of predictors and overall model statistics.


Compare the model to two others
```{r}
models = list(
  model1 = lm(bwt ~ gaweeks + blength, data = birthweight_df),
  model2 = lm(bwt ~ bhead * blength * babysex, data = birthweight_df),
  model3 = lm(bwt ~ fincome + gaweeks + ppwt + ppbmi + malform + smoken, data = birthweight_df))

cv_df =
  crossv_mc(birthweight_df, 100) |> 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble))

cv_df = 
  cv_df |> 
  mutate(
    model1 = map(train, \(df)  lm(bwt ~ gaweeks + blength, data = df)),
    model2 = map(train, \(df)  lm(bwt ~ bhead * blength * babysex, data = df)),
    model3 = map(train, \(df)  lm(bwt ~ fincome + gaweeks + ppwt + ppbmi + malform + smoken, data = df))) |> 
  mutate(
    rmse_model1 = map2_dbl(model1, test, \(mod, df)  rmse(model = mod, data = df)),
    rmse_model2 = map2_dbl(model2, test, \(mod, df)  rmse(model = mod, data = df)),
    rmse_model3 = map2_dbl(model3, test, \(mod, df)   rmse(model = mod, data = df)))

cv_df |> 
  select(starts_with("rmse")) |> 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") |> 
  mutate(model = fct_inorder(model)) |> 
  ggplot(aes(x = model, y = rmse)) + geom_violin()

```
Comment: 
model 1 shows a broad RMSE distribution, predominantly centered around the 350 to 400 range. This suggests a moderate level of prediction error but with some variability across different cross-validation splits.
Model 2 consistently shows the lowest RMSE, suggesting it is the most accurate model for predicting birthweight based on the available predictors.
model 3 has the widest and most varied RMSE distribution, with values ranging significantly from below 300 to above 400. The distribution is somewhat bimodal, suggesting two potential 'modes' of performance depending on the training/test split. This might indicate overfitting or high sensitivity to the specific training data used.
based on the violin plot, Model 2 is recommended as the best choice among the three due to its combination of low and stable RMSE values, indicating good and reliable predictive performance.